---
title: "Intro to Shiny Modules"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(showtext)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#447099",
  secondary_color = "#4D4D4D",
  title_slide_text_color = "#FFFFFF",
  title_slide_background_image = "https://rstudio.com/wp-content/uploads/2018/10/RStudio-Logo-White.png",
  title_slide_background_position = "bottom 20px right 20px",
  title_slide_background_size = "15%",
  background_position = "bottom 20px right 20px",
  inverse_background_color = "#A4C689",
  background_size = "15%",
  background_image = "https://rstudio.com/wp-content/uploads/2018/10/RStudio-Logo-Flat.png",
  text_font_google = google_font("Source Sans Pro"),
  base_font_size = "27px"
)
```
---
class: inverse, center, middle

# Shiny Modules
---
## What are Shiny Modules

A piece of a shiny app, that runs inside a shiny application and can represent input, output or both.

---
class: inverse, center, middle

# 01_why_modules
---

## Why would you use modules?

* Create smaller components of a shiny app that can easily be resused in the same app or in other applications

* Break up complicated apps for easier reasoning & debugging

---
## How?

A shiny modules has two parts:

1) a UI function

2) a piece of server logic using the UI 

---
class: inverse, center, middle

## A simple module example  

```{r eval=FALSE}
counterUI <- function(){}
counterServer <- function(){}
```


---

## Shiny module: UI 
Step 1: Return Shiny UI
```{r eval=FALSE}
counterUI <- function(label = "Counter") {

  tagList(
    actionButton("button", label = label),
    verbatimTextOutput("out")
  )
}
```
---

## Shiny module: UI 

Step 2: Make a Namespace function
```{r eval=FALSE}
counterUI <- function(id, label = "Counter") {
  ns <- NS(id)
  tagList(
    actionButton("button", label = label),
    verbatimTextOutput("out")
  )
}
```
---
## Shiny module: UI 
Step 2: Wrap input and output id's in `ns()`
```{r eval=FALSE}
counterUI <- function(id, label = "Counter") {
  ns <- NS(id)
  tagList(
    actionButton(ns("button"), label = label),
    verbatimTextOutput(ns("out"))
  )
}
```
---

## Shiny module: UI

1. First argument is always `id` 

2. Function body starts with `ns <- NS(id)` to create namespace function

3. All inputs and outputs must be wrapped in `ns()` 

4. Use `tagList` to return multiple UI objects 

---

## Shiny module: Server
Step 1: Add server logic to module server function and be sure toinclude `session` in the server function 
```{r eval=FALSE}
counterServer <- function() {

    function(input, output, session) {
      count <- reactiveVal(0)
      observeEvent(input$button, {
        count(count() + 1)
      })
      output$out <- renderText({
        count()
      })
    }
}
```
---

## Shiny module: Server
Step 2: You don't need to wrap input / outputs in `ns()`
```{r eval=FALSE}
counterServer <- function(id) {

    function(input, output, session) {
      count <- reactiveVal(0)
      observeEvent(input$button, {
        count(count() + 1)
      })
      output$out <- renderText({
        count()
      })
    }

}
```
---

## Shiny module: Server
Step 3: Wrap Server logic function inside `moduleServer()`, pass it an id argument from the function
```{r eval=FALSE}
counterServer <- function(id) {
  moduleServer(
    id,
    function(input, output, session) {
      count <- reactiveVal(0)
      observeEvent(input$button, {
        count(count() + 1)
      })
      output$out <- renderText({
        count()
      })
    }
  )
}
```
---


## Shiny Module: Server 
1. Similar name to corresponding UI function but with server suffix (e.g. `counterUI` and `counterServer`)

2. `id` as input and any additional parameters needed 

3. Calls `moduleServer()`

4. `moduleServer` is passed `id` and a *module function*

5. Module function has `input` `output` and `session` arguments and can access inputs and outputs only from its namespace
---
# Using shiny modules 
```{r eval = FALSE}

ui <- fluidPage(
  counterButton("counter1")
)

server <- function(input, output, session) {
  counterServer("counter1")
}

```

---
## Using shiny modules

1. Call UI function with `id` (and any additional parameters) in ui of shiny app

2. Call server module in the server function of app with matching `id`

3. Repeat above with a unique `id` to reuse the module in the app 

---

## Serving shiny modules to your app 

* *Inline Code:* write module code directly in the app

* *In R Script:* that lives inside the app subdirectory

* *In R Package:* especially if used across other apps 

---
class: inverse, center, middle

# Examples
---

## Nested Modules: UI
```{r eval=FALSE}

innerUI <- function(id) {
  ns <- NS(id)
  "This is the inner UI"
}

outerUI <- function(id) {
  ns <- NS(id)
  wellPanel(
    innerUI(ns("inner1"))
  )
}

```
---
## Nested Modules: Server 

```{r eval=FALSE}
innerServer <- function(id) {
  moduleServer(
    id,
    function(input, output, session) {
    }
  )
}
outerServer <- function(id) {
  moduleServer(
    id,
    function(input, output, session) {
      innerResult <- innerServer("inner1")
    }
  )
}

```

---

# Shiny Module Resources

* [Using Shiny Modules Tutorial](https://rstudio.com/resources/shiny-dev-con/modules/)
* [Effective Use of Shiny Modules](https://rstudio.com/resources/rstudioconf-2019/effective-use-of-shiny-modules-in-application-development/)
* [Modularizing Shiny app code](https://shiny.rstudio.com/articles/modules.html)
